\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-07 paper/paper.tex
Edits to abstract, intro and model assessment
- Fix model 4 plot

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
While reliability is an important aspect of product quality, by its nature, the lifetime data available to consumers is limited since high-reliability consumer products fail infrequently. Due in part to multiple causes of failure, simple parametric models have limited application, being unable to capture relevant lifetime characteristics. On the other hand, more suitable models are difficult to fit because of small sample sizes. If failure analysis data were available it could mitigate identifiability issues; however, consumers do not have this information, nor does it have practical value for them.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout. We fit the GLFP model using a hierarchical Bayesian approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using a Bayesian approach, comparison of different product brands is straightforward as estimation of arbitrary functionals are easily computable from full posterior distributions.
\end{abstract}

@@

\newpage
\tableofcontents
\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Consumers have an interest in accurate assessment of product reliability. Toward this end, there is a need for models that reflect observable facts "on the ground", and methods which can make the most of available data and properly account for uncertainty.
%While computationally convenient, simple lifetime models can fail at each of these.

@@

Nonparametric methods, which provide such flexibility are well-known, Kaplan-Meier for example. However, truncation, censoring and finite samples limit the scope of inference for these methods. Parametric models do not have these limitations. Truncated and/or censored data can be accommodated simply by adjusting the likelihood. Additionally, because they assume a parsimonious representation of the data-generating process, they provide simpler, more intuitive means by which to compare populations of interest, borrow strength across groups, and incorporate prior information.

@@

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling via hierarchical modeling is a way to reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

\subsection{Overview}
The structure of the paper is as follows.  Section~\ref{sec:Background} describes some of the potential applications of our method and summarizes previous related work found in the literature. Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates an application in an example. Section~\ref{sec:Hierarchical GLFP model} introduces a hierarchical model for extending the GLFP model to multiple populations. In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set. Subsection~\ref{sec:Data} describes and summarizes the Backblaze Hard Drive data set.  Subsection~\ref{sec:Model Comparisons} shows the use of approximate leave-one-out cross-validation as a way to do model selection between different hierarchical model specifications. We also address computational issues and modeling decisions that lead to a final model.  In subsection~\ref{sec:Drive-model Comparisons} we present an analysis comparing drive-model lifetime distributions.  Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\cite{betancourt}. Multiple chain were run for 1500 iterations after 1500 warmup iterations. Convergence was assessed by examining posterior
plots and checking that potential scale reduction factors (Rhat) \cite{gelman2014bayesian} were less than $1.1$. Four chains were run for Models 1,2 and 3 and 16 chains were run for Model 4.

@@

To generate replicate data sets, we must first decide what to do with regard to censoring and truncation. The representation we choose to display the data, non-parametric probability plots, guides us in making this decision. The variability of this estimator arises primarily from the number of ``at risk" units at any given time, which depends in part on these processes. Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data. Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}}$ for $s = 1 \dots 19$ is as follows:
\begin{enumerate}
\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.  Where each set of parameters is randomly drawn from the full posterior distribution.
\item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
\item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
\end{enumerate}

We can compare the 19 simulated data sets $y_{rep,g, i}}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator. The resulting plots are displayed in figure \ref{fig:post-pred-KM}. For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data. As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there. Drive-model 4 also appears as an exception, which we discuss now.

@@

To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}). From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year. Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band. Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode.

We can use a similar plot to see how information is borrowed across drive-models. Drive-model 43 is an older model, so that the observed units had run for quite some time before data was made available. In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. Then, in the period where records exist for this drive model, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.

\begin{figure}
\centering
\includegraphics{dm4-exception}
\caption{}
\label{fig:ex-mod-4}
\end{figure}

>>>>

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
When the information carried by lifetime data are limited by sample size, censoring and truncation, lifetime analysis can suffer from imprecision. These limitations are not atypical; for example, lifetime data for high-reliability products will tend to be right-censored.

With field data on consumer products, model-based appproaches to inference require a model with sufficient flexibility to account for multiple kinds of failure. The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using samples from the posterior distribution.
\end{abstract}

@@

\newpage
\tableofcontents
\newpage
\spacingset{1.45} % DON'T change the spacing!
\section{Introduction}
Consumers have an interest in accurate assessment of product reliability. Toward this end, there is a need for models that can accomodate common failure patterns and methods which can make the most of available data and properly account for uncertainty.
%While computationally convenient, simple lifetime models can fail at each of these.

@@

Nonparametric methods require relatively few assumptions, but they are not suitable for forecasting. Moreover, nonparametric estimates made using small samples, or data which are heavily censored may be too noisy to be practical. Appropriate parametric models can make better use of available data. For inference, likelihood-based methods are well-equipped to deal with truncated and/or censored data. Furthermore, by assuming a parsimonious representation of the data-generating process, parametric models admit simpler, more intuitive ways to compare populations of interest, borrow strength across groups, and incorporate prior information.

@@

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

The structure of the paper is as follows.  Section~\ref{sec:Background} describes some of the potential applications of our method and summarizes previous related work found in the literature. Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates an application in an example. Section~\ref{sec:Hierarchical GLFP model} introduces a hierarchical model for extending the GLFP model to multiple populations. In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set. Subsection~\ref{sec:Data} describes and summarizes the Backblaze Hard Drive data set.  Subsection~\ref{sec:Model Comparisons} shows the use of approximate leave-one-out cross-validation as a way to do model selection between different hierarchical model specifications. We also address computational issues and modeling decisions that lead to a final model.  In subsection~\ref{sec:Drive-model Comparisons} we present an analysis comparing drive-model lifetime distributions.  Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan}\cite{rstan} package in {\tt R} \cite{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\cite{betancourt}. HMC jointly updates all model parameters by simulating paths along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4. The Gelman-Rubin potential scale reduction factors, ``R-hat", which test for lack of convergence by comparing the between versus within chain variability were used. Upon convergence, R-hat converges to 1. For the four model fits, the  maximum values over all the model parameters were 1.005, 1.028, 1.009 and 1.004, respectively. The minimum estimated effective sample size over all the model parameters for the four model fits were 567, 205, 672 and 3330, respectively. No divergent transitions occurred after warmup.

% HMC generates proposals by simulating paths along the posterior density. This is done to reduce autocorrelation between samples, but doing so also provides a useful diagnostic. \textit{Divergent transitions} will tend to occur when the posterior has regions of high curvature, which can present problems for MCMC by making it difficult to fully explore the posterior distribution. This feedback can be useful

@@

To generate replicate data sets, we must first decide what to do with regard to censoring and truncation. The representation we choose to display the data, non-parametric probability plots, guides us in making this decision. The variability of this estimator arises primarily from the number of ``at risk" units at any given time, which depends in part on these processes. Since we are not modeling the mechanisms of truncation nor censoring, we want to try to duplicate these mechanisms in the replicated data. Our method for generating a replicate data set  $y_{rep,g,i}^{(s)}$ for $s = 1 \dots 19$ is as follows:
\begin{enumerate}
\item Simulate $\tilde{y}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(s)},\mu_1^{(s)},\sigma_1^{(s)},\mu_{2g}^{(s)}, \sigma_{2g}^{(s)})$.  Where each set of parameters is randomly drawn from the full posterior distribution.
\item If $\tilde{y}_{rep,g,i} < t_{g,i}^L$, then remove this time as this record is not observed
\item Define $m_{g,i}>0$ for each $g$ and $i$. If $c_{g,i}=0$, let $m_{g,i}=\max \{y_{G,i}: G=g\}$, else let $m_{g,i}=y_{g,i}$.\\
If $\tilde{y}_{rep,g,i}^{(s)}>m_{g,i}$, set $y_{rep,g,i}^{(s)}=m_{g,i},\; c_{rep,g,i}^{(s)}=1$, otherwise set $y_{rep,g,i}^{(s)}=\tilde{y}_{rep,g,i}^{(s)},\; c_{rep,g,i}^{(s)}=0$.
\end{enumerate}

We can compare the 19 simulated data sets $y_{rep,g, i}$ to the observed data $y_{g,i}$ by using them to reestimate the Kaplan-Meier nonparametric estimator. The resulting plots are displayed in figure \ref{fig:post-pred-KM}. For most drive-models, the observed estimates appear to behave similarly to those derived from the replicated data. As discussed in \ref{subsec:ex1}, drive-model 14 has a discrepancy in the right tail, which we addressed there. Drive-model 4 also appears as an exception, which we discuss now.

@@

To better understand the reason for the discrepancy between the observed data for drive-model 4 and its posterior predictive distribution, we plot the raw data along with the truncation-adjusted nonparametric estimate of time to failure and the posterior estimate (see figure \ref{fig:ex-mod-4}). From these plots, it is clear that the large discrepancy between the posterior and the nonparametric estimate is driven by a small number of failures that occur in the first quarter year. Indeed, if we ignore records with observation periods that include the first quarter, the modified nonparametric estimate falls within the posterior 90\% pointwise credible band. Our trouble with fitting this drive-model is that while there is evidence of infant mortality in the relatively few units that were observed in the first quarter, the problem seems to go away much quicker than would be expected if there were a large number of units susceptible to the early failure mode. In any case, this example demonstrates that, conditional on the model, the weight of evidence in all the data overpowers the signal from the small sample for this drive-model, which suggests high rates infant mortality in the first quarter year.

For a clear example of this, consider drive-model 43. The data for this drive-model are heavily truncated; the observed units had run for quite some time before data were made available. In figure \ref{fig:ex-mod-43} we see that the lack of data for this drive-model at early ages results in a diffuse posterior centered around the what we call the ``global estimate," corresponding to $\op{GFLP}(\eta_{\pi}, \mu_1,\sigma_1, \eta_{t_2} - \eta_{\sigma_2}\Phi^{-1}(.2), \eta_{\sigma_2})$. Then, in the period where records exist for this drive model, at around $1\frac{1}{2}$ years, the posterior deviates from the global model, acheiving a compromise between it and the nonparametric estimate.

\begin{figure}
\centering
\includegraphics{dm4-exception}
\caption{Top: Graphical representation of lifetime for random sample of 200 drives of drive-model 4 from the Backblaze data set. Solid white lines indicate the observed window of life. Right-censored observations changed to a dashed line and failures mark the transition to a solid black line. Bottom: Solid step function is truncation-adjusted nonparametric estimate of the proportion failing. The dotted line is an estimate obtained after truncating the data to those starting after .25 years, with the corresponding truncation adjustment. The dashed line with gray band is pointwise posterior median, and 90\% credible interval, of proportion failing.}

\label{fig:ex-mod-4}
\end{figure}
